---
title: "SSC Covid Analysis"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo    = TRUE, 
  message = FALSE, 
  warning = FALSE,
  fig.width  = 11,
  fig.height = 5)
```

Load packages.
```{r, include=FALSE}
library(tidyverse)
library(magrittr)
library(covidseir)
library(knitr)
```

Read in data. 
```{r}
dat <- read_csv("../bc_data/BCCDC_COVID19_Dashboard_Case_Details.csv") %>% 
  janitor::clean_names()
dat_daily_cases <- dat %>% 
  group_by(reported_date) %>% 
  summarize(active_cases = n())
```

Plot cases from Jan 1st 2020 to May 15th 2021. We have included noteable BC restrictions in the plot; which we are intersted in learning more about their effect. 
```{r }
major_restrictions <- 
  c("2020-11-07"= "No private gatherings",
    "2020-11-19"= "Masks in stores. \nTravel discouraged.",
    "2021-03-11"= "Outdoor gathering \nof 10 people allowed",
    "2021-03-29"= "No indoor dining\nNo indoor group activity\nMasks grades 4-12")

ggplot(dat_daily_cases, aes(reported_date, active_cases)) + 
  geom_line() + 
  geom_vline(xintercept = as.Date(names(major_restrictions)),
             linetype="dashed",color="darkred", alpha = 0.5 ) +
  geom_text(aes(x=as.Date(names(major_restrictions))[1], 
                label= major_restrictions[1], y=2000), 
            angle = 90, colour="red", text=element_text(size=3)) + 
  geom_text(aes(x=as.Date(names(major_restrictions))[2], 
                label= major_restrictions[2], y=2000), 
            angle = 90, colour="red", text=element_text(size=3)) + 
  geom_text(aes(x=as.Date(names(major_restrictions))[3], 
                label= major_restrictions[3], y=2000), 
            angle = 90, colour="darkgreen", text=element_text(size=3)) +
  geom_text(aes(x=as.Date(names(major_restrictions))[4], 
                label= major_restrictions[4], y=2000), 
            angle = 90, colour="red", text=element_text(size=3)) + 
  ylim(0,3000)
```

For the time being, lets focus on the first pair of restrictions which occured in early Nov 2020. 
```{r}
dat_daily_cases_filt <-  dat_daily_cases %>% 
  filter(reported_date >="2020-10-01") %>% 
  filter(reported_date <= "2021-01-01") %>% 
  mutate(day = 1:nrow(.))

dat_daily_cases_filt_proj <-  dat_daily_cases %>% 
  filter(reported_date >="2020-09-01") %>% 
  mutate(day = 1:nrow(.))

ggplot(dat_daily_cases_filt, aes(reported_date, active_cases)) + 
  geom_line() + 
  geom_vline(xintercept = as.Date(names(major_restrictions)),
             linetype="dashed",color="darkred", alpha = 0.5 ) +
  geom_text(aes(x=as.Date(names(major_restrictions))[1], 
                label= major_restrictions[1], y=2000), 
            angle = 90, colour="red", text=element_text(size=1)) + 
  geom_text(aes(x=as.Date(names(major_restrictions))[2], 
                label= major_restrictions[2], y=2000), 
            angle = 90, colour="red", text=element_text(size=1)) + 
  ylim(0,3000) + 
  scale_x_date(date_breaks = "1 month")

```

Question: Is September 1st to March 1st an appropriate timeline to model the cases influencing/influenced by the early November restrictions?

First model:
```{r}
# First Date/Last Date
first_date = dat_daily_cases_filt$reported_date[1]
last_date = dat_daily_cases_filt$reported_date[nrow(dat_daily_cases_filt)]

# Sample Fraction - estimation of the positive cases that are sampled/detected/
samp_frac <- rep(0.45, nrow(dat_daily_cases_filt))

# Contact rate breakpoints
f_seg_dates <- c(
  first_date, # First day
  as.Date("2020-11-07"), # First restriction
  as.Date("2020-11-19"), # Second restriction
  last_date # Last day
  )

f_seg = c(0)
for ( i in 1:(length(f_seg_dates)-1)) {
  f_seg = c(f_seg, rep(i,f_seg_dates[i+1] - f_seg_dates[i]) )
  }
nrow(dat_daily_cases_filt) == length(f_seg)

# Infected people at initial point in time.
i0 = dat_daily_cases %>% 
  filter((reported_date >= (first_date - 30) & reported_date <= first_date)) %>% 
  pull(active_cases) %>% 
  sum()

# Fit model
fit <- fit_seir(
  daily_cases = dat_daily_cases_filt$active_cases,
  forecast_days = 0,
  samp_frac_fixed = samp_frac, 
  samp_frac_type = "fixed",
  f_seg = f_seg,
  i0_prior = c(log(i0), 1), 
  e_prior = c(0.5, 0.05),
  #start_decline_prior = c(log(30), 0.1),
  #end_decline_prior = c(log(100), 0.1),
  use_ramp = FALSE,
  rw_sigma = 0.2,
  f_prior = cbind(c(0.4, 0.5, 0.6), c(0.2, 0.2, 0.2)),
  R0_prior = c(log(2.6), 0.2), # Does not seem like it has a major effect
  N_pop = 5.15e6, # BC population
  iter = 500, # number of posterior samples
  fit_type = "optimizing" # for speed only
)

proj <- fit %>% 
  covidseir::project_seir(iter = 1:50) 

tidy_proj <- proj %>% 
  covidseir::tidy_seir(resample_y_rep = 20) %>%
  left_join(dat_daily_cases_filt, by = "day")

sum((tidy_proj$active_cases - tidy_proj$y_rep_mean)^2)/nrow(tidy_proj)

# Plot Projection
plot_projection(tidy_proj, 
                obs_dat = dat_daily_cases_filt,
                value_column = "active_cases", date_column = "reported_date"
                )

# Plot Residuals
plot_residuals(tidy_proj, 
               obs_dat = dat_daily_cases_filt, 
               obj = fit, 
               value_column = "active_cases", date_column = "reported_date"
               )

set.seed(1)
plot_residuals(tidy_proj, 
               obs_dat = dat_daily_cases_filt,
               obj = fit, 
               type = "quantile",
               value_column = "active_cases", date_column = "reported_date",
               return_residuals = FALSE
               )

set.seed(1)
resid <- plot_residuals(tidy_proj, 
               obs_dat = dat_daily_cases_filt, 
               obj = fit, 
               type = "quantile",
               value_column = "active_cases", date_column = "reported_date",
               return_residuals = TRUE
               )
hist(resid)
qqline(resid)


```

Projections.
```{r}
days_to_project = dat_daily_cases$reported_date[nrow(dat_daily_cases)] - 
  dat_daily_cases_filt$reported_date[nrow(dat_daily_cases_filt)]
day_start_reduction <- 60

proj2 <- covidseir::project_seir(
  fit,
  iter = 1:50,
  forecast_days = days_to_project,
  f_fixed_start = max(fit$days) + day_start_reduction,
  f_multi = rep(1, days_to_project - day_start_reduction + 1),
  f_multi_seg = 3 # which f segment to use
)

tidy_proj2 <- covidseir::tidy_seir(proj2, resample_y_rep = 30)
tidy_proj2 <- dplyr::left_join(tidy_proj2, dat_daily_cases_filt_proj, by = "day")

covidseir::plot_projection(tidy_proj2, obs_dat = dat_daily_cases_filt_proj,
                           value_column = "active_cases", date_column = "reported_date")
```

