---
title: "Using covidseir to model COVID-19 in B.C. (v3)"
author: "Travis Blimkie, Arjun Baghela & Hossameldin Mohammed"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: readable
    highlight: kate
    fig_caption: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "",
  fig.width  = 9,
  fig.height = 6
)
```


## Load packages
```{r packages_and_setup}
suppressPackageStartupMessages({
  library(covidseir)
  library(lubridate)
  library(tidyverse)
})

theme_set(theme_bw(base_size = 16))
```

<br>

## Load data
We're using the BC COVID-19 case data from the BC CDC, downloaded on May 17th,
2021 from this
[link](http://www.bccdc.ca/health-info/diseases-conditions/covid-19/data).

```{r load_data}
bc_data <-
  read_csv("../bc_data/BCCDC_COVID19_Dashboard_Case_Details.csv") %>%
  janitor::clean_names()

bc_data_daily_cases <- bc_data %>%
  group_by(reported_date) %>%
  summarise(daily_cases = n())
```

<br>

## Plot trends in case data
### Daily cases, January 29 2020 - May 17 2021
```{r daily_cases_all}
ggplot(bc_data_daily_cases, aes(reported_date, daily_cases)) +
  geom_col() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x = "Date",
    y = "Daily Case Count"
  )
```

<br>

### Daily cases, January 29 - December 31 2020
```{r daily_cases_fall}
bc_data_daily_cases %>%
  filter(reported_date <= ymd("20201231")) %>%
  ggplot(., aes(reported_date, daily_cases)) +
  geom_point(pch = 21, fill = "dodgerblue", colour = "black") +
  geom_line() +
  scale_y_continuous(expand = c(0, 20)) +
  labs(
    x = "Date",
    y = "Daily Case Count"
  )
```

<br>

### Daily cases, September 1 - December 31 2020, with highlighted dates
```{r plot_restrictions_labelled}
major_restrictions <- c(
  "2020-11-07" = " Nov. 7th:\nNo private gatherings.",
  "2020-11-19" = " Nov. 19th:\nMasks mandated in stores.\n Non-essential travel\n discouraged."
)

major_restrictions_tbl <- tibble(
  date = names(major_restrictions),
  text = major_restrictions,
  yval = c(175, 380)
)

bc_data_daily_cases %>%
  filter(between(reported_date, ymd("20200901"), ymd("20201231"))) %>%
  ggplot(., aes(reported_date, daily_cases)) +
  geom_point(pch = 21, fill = "dodgerblue", colour = "black", size = 2) +
  geom_line() +
  scale_y_continuous(expand = c(0, 20)) +
  labs(
    x = "Date",
    y = "Daily Case Count"
  ) +
  annotate(
    geom = "vline",
    x = ymd(major_restrictions_tbl$date),
    xintercept = ymd(major_restrictions_tbl$date),
    linetype = "dashed",
    colour = "red",
    size = 1.2
  ) +
  annotate(
    geom = "label",
    x = ymd(major_restrictions_tbl$date),
    y = major_restrictions_tbl$yval,
    label = major_restrictions_tbl$text,
    size = 5,
    vjust = 1,
    hjust = -0.05
  ) +
  theme_bw(base_size = 20)
ggsave("case_plot.png", dpi = 250, height = 6, width = 9)
```

<br>

## Fitting a model to the data
```{r model_seir}
bc_data_fall <- bc_data_daily_cases %>% 
  filter(between(reported_date, ymd(20201001), ymd(20210101))) %>% 
  mutate(
    day = 1:nrow(.),
    f_seg = case_when(
      reported_date == ymd(20201001) ~ 0,
      reported_date <= ymd(20201106) ~ 1,
      reported_date <= ymd(20201118) ~ 2,
      TRUE ~ 3
    )
  )

i0_value <- bc_data_daily_cases %>% 
  filter(between(reported_date, ymd(20201001) - 30, ymd(20201001))) %>% 
  pull(daily_cases) %>% 
  sum()

seir_model_1 <- fit_seir(
  daily_cases     = bc_data_fall$daily_cases,
  forecast_days   = 0,
  samp_frac_fixed = rep(0.45, nrow(bc_data_fall)),
  samp_frac_type  = "fixed",
  f_seg           = bc_data_fall$f_seg,
  i0_prior        = c(log(i0_value), 1), 
  e_prior         = c(0.5, 0.05),
  use_ramp        = FALSE,
  rw_sigma        = 0.2,
  f_prior         = cbind(c(0.4, 0.5, 0.6), c(0.2, 0.2, 0.2)),
  R0_prior        = c(log(2.6), 0.2),
  N_pop           = 5.15e6,
  iter            = 500,
  fit_type        = "optimizing"
)

seir_projection_1 <- project_seir(seir_model_1, iter = 1:50)

seir_tidy_1 <- tidy_seir(seir_projection_1, resample_y_rep = 20) %>% 
  left_join(bc_data_fall, by = "day")

plot_projection(
  pred_dat = seir_tidy_1,
  obs_dat = bc_data_fall,
  value_column = "daily_cases",
  date_column = "reported_date"
) +
  theme_bw(base_size = 18) +
  labs(x = "Date") +
  theme(plot.margin = unit(c(0.5, 0.8, 0.4, 0.4), "cm"))
# ggsave("model_fit_to_data.png", dpi = 250, height = 6, width = 9)


plot_residuals(
  pred_dat = seir_tidy_1,
  obj = seir_model_1,
  obs_dat = bc_data_fall,
  value_column = "daily_cases",
  date_column = "reported_date",
)
```

<br>

## Making projections
### No implementation of restrictions on either date (Nov. 7th or 19th)
```{r project_cases_1}
model_start_date <- ymd(20201001)
model_end_date   <- ymd(20201120)

days_to_project <- ymd(20201215) - model_end_date

bc_data_Nov <- bc_data_daily_cases %>% 
  filter(between(reported_date, model_start_date, model_end_date)) %>% 
  mutate(
    day = 1:nrow(.),
    f_seg = case_when(
      reported_date == model_start_date ~ 0,
      TRUE ~ 1
    )
  )

seir_model_2 <- fit_seir(
  daily_cases     = bc_data_Nov$daily_cases,
  forecast_days   = 0,
  samp_frac_fixed = rep(0.45, nrow(bc_data_Nov)),
  samp_frac_type  = "fixed",
  f_seg           = bc_data_Nov$f_seg,
  i0_prior        = c(log(i0_value), 1), 
  e_prior         = c(0.5, 0.05),
  use_ramp        = FALSE,
  rw_sigma        = 0.2,
  f_prior         = c(0.4, 0.2), # Use first set of values since we have only 1 f_seg
  R0_prior        = c(log(2.6), 0.2),
  N_pop           = 5.15e6,
  iter            = 500,
  fit_type        = "optimizing"
)

seir_projection_2 <- project_seir(
  seir_model_2,
  iter = 1:50,
  forecast_days = days_to_project
)

seir_tidy_2 <- tidy_seir(seir_projection_2, resample_y_rep = 20) %>% 
  left_join(bc_data_Nov, by = "day") %>% 
  mutate(reported_date = seq(model_start_date, ymd(20201215), by = "1 days"))

plot_projection(
  seir_tidy_2,
  obs_dat = bc_data_Nov,
  value_column = "daily_cases",
  date_column = "reported_date"
) +
  theme_bw(base_size = 18) +
  labs(x = "Date", title = "Scenario 1: No implementation of either intervention") +
  theme(plot.margin = unit(c(0.5, 0.9, 0.4, 0.4), "cm"))
# ggsave("model_projection_no_interventions.png", dpi = 250, height = 6, width = 9)
```

<br>

### Restrictions on the 7th, but not the 19th
```{r project_cases_2}
model_start_date <- ymd(20201001)
model_end_date   <- ymd(20201120)

days_to_project <- ymd(20201215) - model_end_date

bc_data_Nov7 <- bc_data_daily_cases %>% 
  filter(between(reported_date, model_start_date, model_end_date)) %>% 
  mutate(
    day = 1:nrow(.),
    f_seg = case_when(
      reported_date == model_start_date ~ 0,
      reported_date <= ymd(20201106) ~ 1,
      TRUE ~ 2
    )
  )

seir_model_3 <- fit_seir(
  daily_cases     = bc_data_Nov7$daily_cases,
  forecast_days   = 0,
  samp_frac_fixed = rep(0.45, nrow(bc_data_Nov7)),
  samp_frac_type  = "fixed",
  f_seg           = bc_data_Nov7$f_seg,
  i0_prior        = c(log(i0_value), 1), 
  e_prior         = c(0.5, 0.05),
  use_ramp        = FALSE,
  rw_sigma        = 0.2,
  f_prior         = cbind(c(0.4, 0.5), c(0.2, 0.2)),
  R0_prior        = c(log(2.6), 0.2),
  N_pop           = 5.15e6,
  iter            = 500,
  fit_type        = "optimizing"
)

seir_projection_3 <- project_seir(
  seir_model_3,
  iter = 1:50,
  forecast_days = days_to_project
)

seir_tidy_3 <- tidy_seir(seir_projection_3, resample_y_rep = 20) %>% 
  left_join(bc_data_Nov7, by = "day") %>% 
  mutate(reported_date = seq(model_start_date, ymd(20201215), by = "1 days"))

plot_projection(
  seir_tidy_3,
  obs_dat = bc_data_Nov7,
  value_column = "daily_cases",
  date_column = "reported_date"
) +
  theme_bw(base_size = 18) +
  labs(x = "Date", title = "Scenario 2: Implementation of Nov. 7th restrictions only") +
  theme(plot.margin = unit(c(0.5, 0.9, 0.4, 0.4), "cm"))
# ggsave("model_projection_intervention_Nov7.png", dpi = 250, height = 6, width = 9)
```

<br>

### Restrictions on the 19th, but not the 7th
```{r project_cases_3}
model_start_date <- ymd(20201001)
model_end_date   <- ymd(20201120)

days_to_project <- ymd(20201215) - model_end_date

bc_data_Nov19 <- bc_data_daily_cases %>% 
  filter(between(reported_date, model_start_date, model_end_date)) %>% 
  mutate(
    day = 1:nrow(.),
    f_seg = case_when(
      reported_date == model_start_date ~ 0,
      reported_date <= ymd(20201118) ~ 1,
      TRUE ~ 2
    )
  )

seir_model_4 <- fit_seir(
  daily_cases     = bc_data_Nov19$daily_cases,
  forecast_days   = 0,
  samp_frac_fixed = rep(0.45, nrow(bc_data_Nov19)),
  samp_frac_type  = "fixed",
  f_seg           = bc_data_Nov19$f_seg,
  i0_prior        = c(log(i0_value), 1), 
  e_prior         = c(0.5, 0.05),
  use_ramp        = FALSE,
  rw_sigma        = 0.2,
  f_prior         = cbind(c(0.4, 0.5), c(0.2, 0.2)),
  R0_prior        = c(log(2.6), 0.2),
  N_pop           = 5.15e6,
  iter            = 500,
  fit_type        = "optimizing"
)

seir_projection_4 <- project_seir(
  seir_model_4,
  iter = 1:50,
  forecast_days = days_to_project
)

seir_tidy_4 <- tidy_seir(seir_projection_4, resample_y_rep = 20) %>% 
  left_join(bc_data_Nov19, by = "day") %>% 
  mutate(reported_date = seq(model_start_date, ymd(20201215), by = "1 days"))

plot_projection(
  seir_tidy_4,
  obs_dat = bc_data_Nov19,
  value_column = "daily_cases",
  date_column = "reported_date"
) +
  theme_bw(base_size = 18) +
  labs(x = "Date", title = "Scenario 3: Implementation of Nov. 19th restrictions only") +
  theme(plot.margin = unit(c(0.5, 0.9, 0.4, 0.4), "cm"))
# ggsave("model_projection_intervention_Nov19.png", dpi = 250, height = 6, width = 9)
```

<br>

### Both interventions 1 week earlier
```{r project_cases_4}
model_start_date <- ymd(20201001)
model_end_date   <- ymd(20201120)

days_to_project <- ymd(20201215) - model_end_date

bc_data_1weekEarly <- bc_data_daily_cases %>% 
  filter(between(reported_date, model_start_date, model_end_date)) %>% 
  mutate(
    day = 1:nrow(.),
    f_seg = case_when(
      reported_date == ymd(20201001) ~ 0,
      reported_date <= ymd(20201030) ~ 1,
      reported_date <= ymd(20201111) ~ 2,
      TRUE ~ 3
    )
  )

seir_model_5 <- fit_seir(
  daily_cases     = bc_data_1weekEarly$daily_cases,
  forecast_days   = 0,
  samp_frac_fixed = rep(0.45, nrow(bc_data_1weekEarly)),
  samp_frac_type  = "fixed",
  f_seg           = bc_data_1weekEarly$f_seg,
  i0_prior        = c(log(i0_value), 1), 
  e_prior         = c(0.5, 0.05),
  use_ramp        = FALSE,
  rw_sigma        = 0.2,
  f_prior         = cbind(c(0.4, 0.5, 0.6), c(0.2, 0.2, 0.2)),
  R0_prior        = c(log(2.6), 0.2),
  N_pop           = 5.15e6,
  iter            = 500,
  fit_type        = "optimizing"
)

seir_projection_5 <- project_seir(
  seir_model_5,
  iter = 1:50,
  forecast_days = days_to_project
)

seir_tidy_5 <- tidy_seir(seir_projection_5, resample_y_rep = 20) %>% 
  left_join(bc_data_1weekEarly, by = "day") %>% 
  mutate(reported_date = seq(model_start_date, ymd(20201215), by = "1 days"))

plot_projection(
  seir_tidy_5,
  obs_dat = bc_data_1weekEarly,
  value_column = "daily_cases",
  date_column = "reported_date"
) +
  theme_bw(base_size = 18) +
  labs(x = "Date", title = "Scenario 4: Both Interventions One Week Earlier") +
  theme(plot.margin = unit(c(0.5, 0.9, 0.4, 0.4), "cm"))
# ggsave("model_projection_bothInterventions1WeekEarly.png", dpi = 250, height = 6, width = 9)
```

<br>

### Both interventions 2 weeks earlier
```{r project_cases_5}
model_start_date <- ymd(20201001)
model_end_date   <- ymd(20201120)

days_to_project <- ymd(20201215) - model_end_date

bc_data_2weeksEarly <- bc_data_daily_cases %>% 
  filter(between(reported_date, model_start_date, model_end_date)) %>% 
  mutate(
    day = 1:nrow(.),
    f_seg = case_when(
      reported_date == ymd(20201001) ~ 0,
      reported_date <= ymd(20201023) ~ 1,
      reported_date <= ymd(20201104) ~ 2,
      TRUE ~ 3
    )
  )

seir_model_6 <- fit_seir(
  daily_cases     = bc_data_2weeksEarly$daily_cases,
  forecast_days   = 0,
  samp_frac_fixed = rep(0.45, nrow(bc_data_2weeksEarly)),
  samp_frac_type  = "fixed",
  f_seg           = bc_data_2weeksEarly$f_seg,
  i0_prior        = c(log(i0_value), 1), 
  e_prior         = c(0.5, 0.05),
  use_ramp        = FALSE,
  rw_sigma        = 0.2,
  f_prior         = cbind(c(0.4, 0.5, 0.6), c(0.2, 0.2, 0.2)),
  R0_prior        = c(log(2.6), 0.2),
  N_pop           = 5.15e6,
  iter            = 500,
  fit_type        = "optimizing"
)

seir_projection_6 <- project_seir(
  seir_model_6,
  iter = 1:50,
  forecast_days = days_to_project
)

seir_tidy_6 <- tidy_seir(seir_projection_6, resample_y_rep = 20) %>% 
  left_join(bc_data_2weeksEarly, by = "day") %>% 
  mutate(reported_date = seq(model_start_date, ymd(20201215), by = "1 days"))

plot_projection(
  seir_tidy_6,
  obs_dat = bc_data_2weeksEarly,
  value_column = "daily_cases",
  date_column = "reported_date"
) +
  theme_bw(base_size = 18) +
  labs(x = "Date", title = "Scenario 5: Both Interventions Two Weeks Earlier") +
  theme(plot.margin = unit(c(0.5, 0.9, 0.4, 0.4), "cm"))
# ggsave("model_projection_bothInterventions2WeeksEarly.png", dpi = 250, height = 6, width = 9)
```

<br>

## Get case numbers at key dates
```{r retrieve_case_nums}
list(
  "No interventions"       = seir_tidy_2, 
  "Nov. 7th intervention"  = seir_tidy_3,
  "Nov. 19th intervention" = seir_tidy_4,
  "Both interventions 1 week early"  = seir_tidy_5,
  "Both interventions 2 weeks early" = seir_tidy_6
) %>% 
  map(~filter(.x, reported_date %in% c(ymd(20201125), ymd(20201201), ymd(20201215)))) %>% 
  bind_rows(., .id = "Model") %>% 
  ungroup() %>% 
  select(Model, reported_date, mu_0.05, mu_0.5, mu_0.95) %>% 
  mutate(
    all_values = paste0(
      round(mu_0.5), " (", round(mu_0.05), "/", round(mu_0.95), ")"
    )
  ) %>% 
  select(-starts_with("mu")) %>% 
  pivot_wider(., names_from = reported_date, values_from = all_values) %>% 
  knitr::kable()
```

<br>

***

## Session Information
```{r sessionInfo}
sessionInfo()
```
