---
title: "Using covidseir to model COVID-19 in B.C."
author: "Travis Blimkie, Arjun Baghela & Hossameldin Mohammed"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: readable
    highlight: kate
    fig_caption: yes
    toc: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo    = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  comment = "",
  fig.width  = 9,
  fig.height = 6
)
```


## Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(covidseir)
  library(lubridate)
  library(tidyverse)
  # future::plan(future::multisession)
})
```

<br>

## Load data
We're using the BC COVID-19 case data from the BC CDC, downloaded on May 17th,
2021 from this
[link](http://www.bccdc.ca/health-info/diseases-conditions/covid-19/data).

```{r load_data}
bc_data <-
  read_csv("../bc_data/BCCDC_COVID19_Dashboard_Case_Details.csv") %>%
  janitor::clean_names()

bc_data_daily_cases <- bc_data %>%
  group_by(reported_date) %>%
  summarise(daily_cases = n())
```

<br>

## Plot trends in case data
### Daily Cases, January 29 2020 - May 17 2021
```{r daily_cases_all}
ggplot(bc_data_daily_cases, aes(reported_date, daily_cases)) +
  geom_col() +
  theme_bw(base_size = 16) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x = "Date",
    y = "Daily Case Count"
  )
```

<br>

### Daily cases, January 29 2020 - May 17 2021, coloured by region
```{r daily_cases_region}
bc_data %>% 
  group_by(reported_date, ha) %>% 
  summarise(daily_cases = n()) %>% 
  ggplot(., aes(reported_date, daily_cases, fill = ha)) +
  geom_col(alpha = 0.5) +
  theme_bw(base_size = 16) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x = "Date",
    y = "Daily Case Count",
    fill = "Region"
  )
```

<br>

### Daily cases, January 29 - December 31 2020
```{r daily_cases_fall}
bc_data_daily_cases %>%
  filter(reported_date <= ymd("20201231")) %>%
  ggplot(., aes(reported_date, daily_cases)) +
  geom_point(pch = 21, fill = "dodgerblue", colour = "black") +
  geom_line() +
  theme_bw(base_size = 16) +
  scale_y_continuous(expand = c(0, 20)) +
  labs(
    x = "Date",
    y = "Daily Case Count"
  )
```

<br>

### Daily cases, September 1 - December 31 2020, with highlighted dates
```{r plot_restrictions_labelled}
major_restrictions <- c(
  "2020-11-07" = " Nov. 7th:\nNo private gatherings.",
  "2020-11-19" = " Nov. 19th:\nMasks mandated in stores.\n Non-essential travel\n discouraged."
  # "2021-03-11" = "Outdoor gathering \nof 10 people allowed",
  # "2021-03-29" = "\n\n\nNo indoor dining\nNo indoor group activity\nMasks grades 4-12"
)

major_restrictions_tbl <- tibble(
  date = names(major_restrictions),
  text = major_restrictions,
  yval = c(200, 400)
)

bc_data_daily_cases %>%
  filter(between(reported_date, ymd("20200901"), ymd("20201231"))) %>%
  ggplot(., aes(reported_date, daily_cases)) +
  geom_point(pch = 21, fill = "dodgerblue", colour = "black") +
  geom_line() +
  theme_bw(base_size = 16) +
  scale_y_continuous(expand = c(0, 20)) +
  labs(
    x = "Date",
    y = "Daily Case Count"
  ) +
  annotate(
    geom = "vline",
    x = ymd(major_restrictions_tbl$date),
    xintercept = ymd(major_restrictions_tbl$date),
    linetype = "dashed",
    colour = "red",
    size = 1
  ) +
  annotate(
    geom = "label",
    x = ymd(major_restrictions_tbl$date),
    y = major_restrictions_tbl$yval,
    label = major_restrictions_tbl$text,
    vjust = 1,
    hjust = -0.05
  )
```

<br>

## The starting-point model
Our current model spans from March 2020 to March 2021. It seems to perform 
reasonably well the in beginning, but the residuals quickly become quite large
once cases numbers start to rise and then peak in late 2020.
```{r current_model}
model_input_v0 <- bc_data_daily_cases %>%
  filter(reported_date < ymd("20210315")) %>% 
  mutate(day = 1:nrow(.))

samp_frac_v0 <- c(
  rep(0.14, 13),
  rep(0.21, 38),
  rep(0.37, 125),
  rep(0.45, nrow(model_input_v0) - (13 + 38 + 125))
)

model_input_v0 <- model_input_v0 %>%
  mutate(
    f_seg = case_when(
      reported_date == ymd("2020-01-29") ~ 0,
      reported_date < ymd("20200501") ~ 1,
      between(reported_date, ymd("2020-05-01"), ymd("2020-06-01")) ~ 2,
      between(reported_date, ymd("2020-06-02"), ymd("2020-11-15")) ~ 3,
      between(reported_date, ymd("2020-11-16"), ymd("2021-01-01")) ~ 4,
      between(reported_date, ymd("2021-01-02"), ymd("2021-05-05")) ~ 5,
      TRUE ~ 9
    )
  )

seir_real_dat_v0 <- fit_seir(
  daily_cases         = model_input_v0$daily_cases,
  f_seg               = model_input_v0$f_seg,
  samp_frac_fixed     = samp_frac_v0,
  days_back           = 60,
  i0_prior            = c(log(8), 1),
  e_prior             = c(0.8, 0.05),
  start_decline_prior = c(log(15), 0.1),
  end_decline_prior   = c(log(22), 0.1),
  f_prior             = rbind(c(0.4, 0.2), c(0.5, 0.2), c(0.6, 0.2), c(0.7, 0.2), c(0.7, 0.2)),
  R0_prior            = c(log(2.95), 0.07),
  N_pop               = 5.1e6,
  fit_type            = "optimizing",
  iter                = 50
)

project_real_dat_v0 <- project_seir(seir_real_dat_v0, iter = 1:50)

project_real_tidy_v0 <- 
  tidy_seir(project_real_dat_v0, resample_y_rep = 20) %>%
  left_join(., model_input_v0, by = "day")

plot_projection(
  pred_dat     = project_real_tidy_v0,
  obs_dat      = model_input_v0,
  value_column = "daily_cases",
  date_column  = "reported_date",
) +
  theme_bw(base_size = 16) +
  labs(x = "Date")

plot_residuals(
  pred_dat      = project_real_tidy_v0,
  obs_dat       = model_input_v0,
  obj           = seir_real_dat_v0,
  value_column  = "daily_cases",
  date_column   = "day",
  type          = c("raw", "quantile"),
  date_function = lubridate::ymd
) +
  theme_bw(base_size = 16) +
  labs(x = "Day")
```

<br>

## Research question
In November 2020, the public health authority introduced new measures in an
attempt to slow the spread of COVID-19 in the province. These included a
mandatory mask policy for all indoor public spaces, travel restrictions and the
banning of private gatherings with individuals outside the household. Can we
develop a model based on BC's COVID data to project possible outcomes had these
measures been implemented earlier or later?

<br>

## Developing a new model

### Parameters of concern
Parameters we need to figure out, in approximate order of importance/priority:

- R0 - I think this should be estimated from our case data for the time range
we're interested in, but I'm not sure how to do this...
- Sampling fraction - the `covidseir` package covered the first 176 days of the
model, and I covered the remaining days with a value of 0.45, thinking testing
coverage probably increased somewhat...but this number needs to be (ideally)
based on some sort of data or public health record
- Distancing dates - these "breakpoints" represent changes in social distancing
behaviour, and their number and placement in the time line need to be adjusted
- Days back - needs to be "...sufficiently large tha the results don't change,"
from the `fit_seir` help page. I don't know what the default is, but I set it to
60 in the above models.
- The social distancing ramp (start- and end-decline prior) could also be
tweaked, though I used the values from the vignette which is based on BC data;
It is from a shorter time range than what we're looking at, though it should
still be applicable since we're just extending the end date of the model

### Attempt 1
Since at moment I don't want to try and figure out how to ignore the social
distancing ramp (to allow us to start at a later date, e.g. September), let's 
just filter the data to the end of December 2020. And I also reverted to using 
the vignette's estimate of R0 in B.C., since at the moment I can't justify 
using a different value.
```{r attempt1}
model_input_v1 <- bc_data_daily_cases %>%
  filter(reported_date <= ymd("20201231")) %>% 
  mutate(day = 1:nrow(.))

samp_frac_v1 <- c(
  rep(0.14, 13),
  rep(0.21, 38),
  rep(0.37, 125),
  rep(0.45, nrow(model_input_v1) - (13 + 38 + 125))
)

model_input_v1 <- model_input_v1 %>%
  mutate(
    f_seg = case_when(
      reported_date == ymd("2020-01-29") ~ 0,
      reported_date < ymd("20200501") ~ 1,
      between(reported_date, ymd("2020-05-01"), ymd("2020-06-01")) ~ 2,
      between(reported_date, ymd("2020-06-02"), ymd("2020-11-15")) ~ 3,
      between(reported_date, ymd("2020-11-16"), ymd("2020-12-31")) ~ 4,
      TRUE ~ 9
    )
  )

seir_real_dat_v1 <- fit_seir(
  daily_cases         = model_input_v1$daily_cases,
  f_seg               = model_input_v1$f_seg,
  samp_frac_fixed     = samp_frac_v1,
  days_back           = 60,
  i0_prior            = c(log(8), 1),
  e_prior             = c(0.8, 0.05),
  start_decline_prior = c(log(15), 0.1),
  end_decline_prior   = c(log(22), 0.1),
  f_prior             = rbind(c(0.4, 0.2), c(0.5, 0.2), c(0.6, 0.2), c(0.7, 0.2)),
  R0_prior            = c(log(2.6), 0.2),
  N_pop               = 5.1e6,
  fit_type            = "optimizing",
  iter                = 50
)

project_real_dat_v1 <- project_seir(seir_real_dat_v1, iter = 1:50)

project_real_tidy_v1 <- 
  tidy_seir(project_real_dat_v1, resample_y_rep = 20) %>%
  left_join(., model_input_v1, by = "day")

plot_projection(
  pred_dat     = project_real_tidy_v1,
  obs_dat      = model_input_v1,
  value_column = "daily_cases",
  date_column  = "reported_date",
) +
  theme_bw(base_size = 16) +
  labs(x = "Date")

plot_residuals(
  pred_dat      = project_real_tidy_v1,
  obs_dat       = model_input_v1,
  obj           = seir_real_dat_v1,
  value_column  = "daily_cases",
  date_column   = "day",
  type          = c("raw", "quantile"),
  date_function = lubridate::ymd
) +
  theme_bw(base_size = 16) +
  labs(x = "Day")
```

<br>

So we still have a problem with the residuals going crazy along with our spike
in cases over the fall (BC's second wave)...

<br>

### Attempt 2
It seems the previous estimate of R0 I was using (2.95) came from the
supplemental material of a publication from the `covidseir` group. Here's the
[link](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1008274#sec008);
I'm going to start using that value again and tweak some other parameters.
```{r attempt2}
model_input_v2 <- bc_data_daily_cases %>%
  filter(reported_date <= ymd("20201231")) %>% 
  mutate(day = 1:nrow(.))

model_input_v2 <- model_input_v2 %>%
  mutate(
    f_seg = case_when(
      reported_date == ymd("20200129") ~ 0,
      reported_date <= ymd("20200430") ~ 1,
      reported_date <= ymd("20200531") ~ 2,
      reported_date <= ymd("20201114") ~ 3,
      reported_date <= ymd("20201231") ~ 4,
      TRUE ~ 9
    ),
    samp_frac = case_when(
      reported_date <= ymd("20200311") ~ 0.14,
      reported_date <= ymd("20200418") ~ 0.21,
      reported_date <= ymd("20200831") ~ 0.37,
      reported_date >= ymd("20200901") ~ 0.45,
      TRUE ~ 9
    )
  )

seir_real_dat_v2 <- fit_seir(
  daily_cases         = model_input_v2$daily_cases,
  f_seg               = model_input_v2$f_seg,
  samp_frac_fixed     = model_input_v2$samp_frac,
  days_back           = 60,
  i0_prior            = c(log(8), 1),
  e_prior             = c(0.8, 0.05),
  start_decline_prior = c(log(15), 0.1),
  end_decline_prior   = c(log(22), 0.1),
  f_prior             = rbind(c(0.4, 0.2), c(0.5, 0.2), c(0.6, 0.2), c(0.6, 0.2)),
  R0_prior            = c(log(2.95), 0.07),
  N_pop               = 5.1e6,
  fit_type            = "optimizing",
  iter                = 50
)

project_real_dat_v2 <- project_seir(seir_real_dat_v2, iter = 1:50)

project_real_tidy_v2 <- 
  tidy_seir(project_real_dat_v2, resample_y_rep = 20) %>%
  left_join(., model_input_v2, by = "day")

plot_projection(
  pred_dat     = project_real_tidy_v2,
  obs_dat      = model_input_v2,
  value_column = "daily_cases",
  date_column  = "reported_date",
) +
  theme_bw(base_size = 16) +
  labs(x = "Date")

plot_residuals(
  pred_dat      = project_real_tidy_v2,
  obs_dat       = model_input_v2,
  obj           = seir_real_dat_v2,
  value_column  = "daily_cases",
  date_column   = "day",
  type          = c("raw", "quantile"),
  date_function = lubridate::ymd
) +
  theme_bw(base_size = 16) +
  labs(x = "Day")
```

<br<

## Trying some projections
```{r projection1}
model_input_v3 <- bc_data_daily_cases %>%
  filter(reported_date <= ymd("20201115")) %>% 
  mutate(day = 1:nrow(.))

all_days_v3 <- bc_data_daily_cases %>% 
  filter(reported_date <= ymd("20201231")) %>% 
  mutate(day = 1:nrow(.)) %>% 
  select(day, reported_date)

model_input_v3 <- model_input_v3 %>%
  mutate(
    f_seg = case_when(
      reported_date == ymd("20200129") ~ 0,
      reported_date <= ymd("20200430") ~ 1,
      reported_date <= ymd("20200531") ~ 2,
      reported_date <= ymd("20201115") ~ 3,
      TRUE ~ 9
    ),
    samp_frac = case_when(
      reported_date <= ymd("20200311") ~ 0.14,
      reported_date <= ymd("20200418") ~ 0.21,
      reported_date <= ymd("20200831") ~ 0.37,
      reported_date >= ymd("20200901") ~ 0.45,
      TRUE ~ 9
    )
  )

seir_real_dat_v3 <- fit_seir(
  daily_cases         = model_input_v3$daily_cases,
  f_seg               = model_input_v3$f_seg,
  samp_frac_fixed     = model_input_v3$samp_frac,
  days_back           = 60,
  i0_prior            = c(log(8), 1),
  e_prior             = c(0.8, 0.05),
  start_decline_prior = c(log(15), 0.1),
  end_decline_prior   = c(log(22), 0.1),
  f_prior             = rbind(c(0.4, 0.2), c(0.5, 0.2), c(0.6, 0.2)),
  R0_prior            = c(log(2.95), 0.07),
  N_pop               = 5.1e6,
  fit_type            = "optimizing",
  iter                = 50
)

project_real_dat_v3 <- project_seir(
  obj           = seir_real_dat_v3, 
  iter          = 1:50, 
  forecast_days = 45,
  f_fixed_start = max(seir_real_dat_v3$days) + 5,
  f_multi       = rep(0.67, 45 - 5 + 1),
  f_multi_seg   = 3
)

project_real_tidy_v3 <- 
  tidy_seir(project_real_dat_v3, resample_y_rep = 20) %>% 
  left_join(., all_days_v3, by = "day")

plot_projection(
  pred_dat     = project_real_tidy_v3,
  obs_dat      = model_input_v3,
  value_column = "daily_cases",
  date_column  = "reported_date",
) +
  theme_bw(base_size = 16) +
  labs(x = "Date")
```

<br>

***

```{r session_info}
sessionInfo()
```

